---
title: "Population genetics of *Melampsora larici-populina* based on microsatellite markers"
subtitle: "Analysis workflow for *Long-lasting coexistence of multiple asexual lineages alongside their sexual counterparts in a fungal plant pathogen*"
author: "Ammar Abdalrahem"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: cerulean
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    number_sections: true
    code_folding: show
---

Section numbers and the table of contents are generated automatically from the
headings below, so they stay correct when sections are added or moved. Headings
name the figure or table of the manuscript that the section produces.




# Setup and data

## Install and load required packages

```{r, include=FALSE}
# Install (if necessary) and load required packages.
# Packages are installed in dependency-ordered layers so that prerequisites
# (e.g. sf, terra, units) are built before the packages that depend on them.
# Installing everything in a single call can fail because packages with
# heavy system dependencies start building before their prerequisites exist.

options(repos = c(CRAN = "https://cloud.r-project.org"))

install_if_missing <- function(pkgs, bioc = FALSE) {
  missing <- pkgs[!pkgs %in% rownames(installed.packages())]
  if (!length(missing)) return(invisible(NULL))
  message("Installing: ", paste(missing, collapse = ", "))
  if (bioc) {
    BiocManager::install(missing, ask = FALSE, update = FALSE, Ncpus = 1)
  } else {
    install.packages(missing, Ncpus = 1)
  }
}

# 1. CRAN core
cran_core <- c("knitr", "ggplot2", "readxl", "tidyverse", "cowplot",
               "viridis", "ggrepel", "ggsci", "scales", "dplyr",
               "factoextra", "grid", "svglite", "reshape2")
install_if_missing(cran_core)

# 2. Spatial stack
cran_spatial <- c("sf", "rnaturalearth", "rnaturalearthdata",
                  "mapdata", "mapplots")
install_if_missing(cran_spatial)
# rnaturalearthhires is hosted on r-universe (not CRAN); required for scale = "large"
if (!"rnaturalearthhires" %in% rownames(installed.packages())) {
  install.packages("rnaturalearthhires",
                   repos = "https://ropensci.r-universe.dev")
}

# 3. Genetics / bio (CRAN)
cran_bio <- c("lme4", "genepop", "hierfstat", "adegenet",
              "poppr", "pegas", "ape", "ade4")
install_if_missing(cran_bio)

# 4. Bioconductor setup (version locked to match R 4.4.x)
if (!requireNamespace("BiocManager", quietly = TRUE)) {
  install.packages("BiocManager")
}
BiocManager::install(version = "3.20", ask = FALSE, update = FALSE)

# 5. Bioconductor packages
install_if_missing(c("ggtree", "ggtreeExtra"), bioc = TRUE)

# 6. GitHub package: RClone
if (!"RClone" %in% rownames(installed.packages())) {
  if (!requireNamespace("devtools", quietly = TRUE)) install.packages("devtools")
  tryCatch(
    devtools::install_github("dbailleul/RClone", dependencies = FALSE),
    error = function(e) message("RClone install failed: ", e$message)
  )
}

# 7. Load all packages safely
all_pkgs <- c(cran_core, cran_spatial, "rnaturalearthhires", cran_bio,
              "ggtree", "ggtreeExtra", "RClone")
invisible(lapply(all_pkgs, function(p) {
  if (requireNamespace(p, quietly = TRUE)) {
    suppressPackageStartupMessages(library(p, character.only = TRUE))
  } else {
    message("Skipping missing package: ", p)
  }
}))

# 8. Shared colour palette
# The NPG (Nature Publishing Group) palette is used by Figure 2, Figure 3,
# Figure 4 and Figure S4. It is defined once here, with the packages, so that
# no figure chunk depends on another figure chunk having been run first.
npg_colors <- pal_npg("nrc", alpha = 0.9)(9)

```



## Data retrieval

```{r, include=FALSE}
# Set working directory to the Rmd's own directory.
# When knitting, knitr already uses the Rmd directory; this block handles
# interactive chunk execution inside RStudio as well.
if (!isTRUE(getOption("knitr.in.progress"))) {
  if (requireNamespace("rstudioapi", quietly = TRUE) && rstudioapi::isAvailable()) {
    setwd(dirname(rstudioapi::getSourceEditorContext()$path))
  }
}

#print the path
getwd()

# Create output sub-directories (figures and tables)
dir.create("output/figures", recursive = TRUE, showWarnings = FALSE)
dir.create("output/tables",  recursive = TRUE, showWarnings = FALSE)


# Import data from CSV
general_data <- read_csv("Table_S1.csv", show_col_types = FALSE)

# covert data to data frame
general_data <- as.data.frame(general_data) 

# remove unknown population
general_data <- general_data[!(general_data$Pop == "NA"), ]
# remove the degree symbol from all character columns in the data frame
general_data <- general_data %>%
  mutate(across(where(is.character), ~ gsub("°", "", .)))

# Standardise Dept names: use Dept # as the authoritative key.
# Catches known variants (missing hyphens, missing accents) that have been
# observed in different versions of the dataset.
dept_canonical <- c(
  "04" = "Alpes-de-Haute-Provence",  # also seen as 'Alpes de Haute-Provence'
  "34" = "Hérault",                  # also seen as 'Herault' (missing accent)
  "54" = "Meurthe-et-Moselle"        # also seen as 'Meurthe-et Moselle'
)
fix_mask <- general_data[["Dept #"]] %in% names(dept_canonical)
general_data$Dept[fix_mask] <- dept_canonical[general_data[["Dept #"]][fix_mask]]




```

```{r, echo=FALSE}
head(general_data)
```

## Figure S1 - sampling sites and sampling strategy

```{r}

# 1) Summarise isolates per site/year
sampling_summary <- general_data %>%
  group_by(Project, Site, Dept, Long, Lat, Year) %>%
  summarise(n_isolates = n(), .groups = "drop") %>%
  mutate(
    Campaign = case_when(
      Project == "Fr09" ~ "National survey 2009",
      Project == "Fr11" ~ "National survey 2011",
      Project %in% c("Dur04", "Dur08", "Dur20") ~ "Durance River valley monitoring",
      Year == 2024 & grepl("Cavalaire", Site, ignore.case = TRUE) ~ "Cavalaire-sur-Mer (Feb 2024)",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(Campaign))

# Split: circles vs star
pts <- sampling_summary %>% filter(Campaign != "Cavalaire-sur-Mer (Feb 2024)")
cav <- sampling_summary %>%
  filter(Campaign == "Cavalaire-sur-Mer (Feb 2024)") %>%
  slice(1) %>%
  mutate(StarLegend = "Cavalaire-sur-Mer (Feb 2024)")

# 2) Map layers
fr <- ne_countries(scale = "large", country = "France", returnclass = "sf")
fr_depts <- ne_states(country = "France", returnclass = "sf")

# 3) Palette
cols <- c(
  "National survey 2009" = "#2C7BB6",
  "National survey 2011" = "#F29E4C",
  "Durance River valley monitoring" = "#2A9D8F"
)

# 4) Plot
p <- ggplot() +
  geom_sf(data = fr, fill = "white", color = "black", linewidth = 0.35) +
  geom_sf(data = fr_depts, fill = NA, color = "grey85", linewidth = 0.20) +

  # Durance highlight
  annotate("rect", xmin = 5.7, xmax = 6.7, ymin = 43.55, ymax = 44.15,
           fill = "#2A9D8F", alpha = 0.18, color = NA) +

  # Circles
  geom_point(
    data = pts,
    aes(Long, Lat, fill = Campaign, size = n_isolates),
    shape = 21, color = "black", stroke = 0.25, alpha = 0.95
  ) +

  # Cavalaire star + legend
  geom_point(
    data = cav,
    aes(Long, Lat, shape = StarLegend),
    color = "#B2182B", size = 4.8, stroke = 1
  ) +

  # Durance label + arrow
  annotate("text", x = 7.35, y = 44.55, label = "Durance\nRiver valley",
           size = 4.2, hjust = 0) +
  annotate("segment", x = 7.20, xend = 6.35, y = 44.40, yend = 43.90,
           arrow = arrow(length = unit(2, "mm")), linewidth = 0.6) +

  # Cavalaire label + arrow (arrow points to the star coordinates)
  annotate("text", x = 6.5, y = 42.85, label = "Cavalaire-\nsur-Mer",
           size = 4.2, hjust = 0) +
  annotate("segment",
           x = 6.45, y = 42.78,
           xend = cav$Long[1], yend = cav$Lat[1],
           arrow = arrow(length = unit(2, "mm")),
           linewidth = 0.6) +

  # Scales
  scale_fill_manual(values = cols, name = NULL) +
  scale_shape_manual(values = c("Cavalaire-sur-Mer (Feb 2024)" = 8), name = NULL) +  # <-- FIX
  scale_size_continuous(
    name = "Size scale",
    range = c(1.5, 7.0),
    breaks = c(5, 20, 50),
    labels = c("5 isolates", "20 isolates", "50+ isolates")
  ) +

  coord_sf(xlim = c(-5, 9.7), ylim = c(41.3, 51.6), expand = FALSE) +

  guides(
    size  = guide_legend(order = 1, override.aes = list(fill = "white", shape = 21, color = "black")),
    fill  = guide_legend(order = 2, override.aes = list(size = 3, shape = 21, color = "black")),
    shape = guide_legend(order = 3, override.aes = list(size = 4, color = "#B2182B"))
  ) +

  theme_void(base_size = 12) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8),
    legend.position = "right",
    legend.box = "vertical",
    legend.spacing.y = unit(3, "mm"),
    legend.text  = element_text(size = 12),
    legend.title = element_text(size = 12),
    plot.margin = margin(6, 40, 6, 6),
    plot.title = element_text(size = 14)
  ) +
  labs(title = "Geographic distribution map")

p


# Optional editable SVG (often very convenient in Inkscape)
# install.packages("svglite")
ggsave("output/figures/FigS1_geographic_distribution.svg", p, device = svglite::svglite,
       width = 250, height = 190, units = "mm")




```

# Delineation of multilocus lineages (manuscript Results 3.1)

## From multilocus genotypes (MLG) to multilocus lineages (MLL)

```{r}
# Create table of genotype data by select columns with "Mlp" in the name and the first column as isolate id
genotype_cols <- c("Isolate", grep("Mlp", names(general_data), value = TRUE)) 
genotype_data <- general_data[, genotype_cols] # 

# Make isolate id as column names 
rownames(genotype_data) <- genotype_data[,1]
# delete the first column not needed anymore
genotype_data$Isolate = NULL

# Remove all non-ASCII characters from individual names
rownames(genotype_data) <- iconv(rownames(genotype_data), 
                                  from = "UTF-8", to = "ASCII", sub = "")



# Convert data to genind object
data_GenInd <- df2genind(
  X = genotype_data,   
  sep = NULL,
  ncode = 3,           # the number of characters used for coding one genotype at one locus.
  ind.names = rownames(genotype_data),  
  loc.names = colnames(genotype_data),  
  pop = general_data$Pop,               # giving the population of each individual
  NA.char = "999",                      # string corresponding to missing allele 999 or 999999
  ploidy = 2,
  type = "codom",                       
  strata = NULL,      
  hierarchy = NULL,
  check.ploidy = getOption("adegenet.check.ploidy") 
)

# Convert data to genclone object
data_Genclone <- as.genclone(data_GenInd)

# Define MLG according threshold = 0
# Means that two individuals are considered to be from the same MLG if their genetic distance is 0
# The "asis" option is used to indicate that missing data should be treated "as is," meaning that missing values will be retained as NA in the distance matrix.
mlg_assignments <- mlg.filter(data_Genclone, threshold = 0, distance = "diss.dist", threads = 1L,missing = "asis") 


# Add MLG result to the table
genotype_data$MLG <- mlg_assignments

#Define the MLL 
#choose a threshold to represent the minimum genetic distance 
#grouping individuals into the same MLL

data_filtered <-filter_stats(data_Genclone, distance = diss.dist, plot = TRUE, missing = "asis",hist= "sturges")

# “farthest neighbor” algorithm.
#Arnaud-Haond et al. 2007, @bailleul2016rclone
print(farthest_thresh <- cutoff_predictor(data_filtered$farthest$THRESHOLDS))

# “UPGMA ” algorithm.
print(average_thresh  <- cutoff_predictor(data_filtered$average$THRESHOLDS))

# “nearest neighbor” algorithm.
print(nearest_thresh  <- cutoff_predictor(data_filtered$nearest$THRESHOLDS))

#Define the MLL threshold 
#The threshold is set to X to represent the minimum genetic distance for grouping individuals into the same MLL

mll_assignments<- mlg.filter(data_Genclone, threshold = average_thresh , algorithm = "average_neighbor" , distance = "diss.dist", threads = 1L, missing = "asis" ) 

genotype_data$MLL <- mll_assignments

# EXPORT: numbers quoted in Results 3.1 (number of MLGs, number of MLLs and the
# genetic distance threshold used for MLL delineation) had no output file.
mll_delineation <- data.frame(
  N_individuals      = nrow(genotype_data),
  N_MLG              = length(unique(genotype_data$MLG)),
  N_MLL              = length(unique(genotype_data$MLL)),
  threshold_UPGMA    = average_thresh,
  threshold_farthest = farthest_thresh,
  threshold_nearest  = nearest_thresh
)
kable(mll_delineation, digits = 3)
write.csv(mll_delineation,
          file = "output/tables/Table_MLL_delineation_summary.csv",
          row.names = FALSE)

```

## Ordering multilocus lineages by abundance


```{r}
# Calculate the frequency of each MLL combination
count_MLL_table <- data.frame(MLL = unique(genotype_data$MLL), Count = NA)

# For each MLL value, count the number of occurrences in the genotype data
# then store the count in the "Count" column of the count_MLL_table
for (i in 1:nrow(count_MLL_table)) {
  mll_value <- count_MLL_table$MLL[i]
  count_MLL_table$Count[i] <- sum(genotype_data$MLL == mll_value)
}


count_MLL_table <- count_MLL_table[order(-count_MLL_table$Count), ]


count_MLL_table$N <- 1:nrow(count_MLL_table)

# Add row names to count_MLL_table
rownames(count_MLL_table) <- NULL


genotype_data$temp_row_names <- rownames(genotype_data)


genotype_data <- merge(genotype_data, count_MLL_table, by = "MLL", all.x = TRUE)

# Restore the original row names
rownames(genotype_data) <- genotype_data$temp_row_names

# Remove the temporary column
genotype_data$temp_row_names <- NULL


colnames(genotype_data)[colnames(genotype_data) == "N"] <- "organised_MLL"


# Ascending order 
genotype_data <- genotype_data[order(genotype_data$organised_MLL),]

```


# Reproductive mode assessment (manuscript Results 3.2)

## Clustering approach

### Assignment of isolates to clusters - Figures S2 and S3

```{r}
# Create a table of genotypes for allele data only
# Select columns with "Mlp" in the name and the first column as isolate id
genotype_cols <- c(grep("Mlp", names(genotype_data), value = TRUE))
genotype_mlp <- genotype_data[, genotype_cols]


# Make isolate id as column names 
rownames(genotype_mlp) <- rownames(genotype_data)

# Convert data to genind object
data_GenInd <- df2genind(
  X = genotype_mlp,   # data.frame containing allele data 
  sep = NULL,
  ncode = 3,           # the number of characters used for coding one genotype at one locus.
  ind.names = rownames(genotype_mlp),  
  loc.names = colnames(genotype_mlp),  
  pop = genotype_data$MLL,               # giving the mll of each individual
  NA.char = "999",                      # string corresponding to missing allele 999 or 999999
  ploidy = 2,
  type = "codom",                       
  strata = NULL,      
  hierarchy = NULL,
  check.ploidy = getOption("adegenet.check.ploidy") 
)


# choose the best K
X <- adegenet::tab(data_GenInd, freq = TRUE, NA.method = "mean")

# 2) PCA to reduce dimension
set.seed(123)

pca <- prcomp(X, center = TRUE, scale. = FALSE)
opt_pcs <- 10  # temporary fixed number of PCs to retain for clustering (can be optimized with xvalDapc later)
scores <- pca$x[, 1:opt_pcs, drop = FALSE]

# 3) Silhouette-based K selection on PCA scores
# Define custom kmeans with higher iterations
my_kmeans <- function(x, k) {
  kmeans(x, k, iter.max = 100, nstart = 25)
}

# Figure S2 - silhouette score for k = 1..20.
# The plot is computed once and stored. The previous version called
# fviz_nbclust twice (once on screen, once inside the png device), which ran
# the same slow k-means twice, and a ggplot object is not auto-printed inside
# a device, so the saved png could come out empty. print() fixes that.
sil_plot <- fviz_nbclust(scores, my_kmeans, method = "silhouette", k.max = 20)
sil_plot

# save the plot
png("output/figures/FigS2_Silhouette_kmeans.png", width = 1200, height = 800, res = 150)
print(sil_plot)
dev.off()

# 4) find.clusters()

grp <- find.clusters(data_GenInd, method = "kmeans", stat = "BIC", n.pca= 90 , n.clust= 2, n.iter=10000, n.start=100)
cluster_assignments <- grp$grp # Assign the cluster assignments to a variable
genotype_data$cluster <- cluster_assignments # Add the cluster assignments to the genotype data table


# 5) Cross-validation to find optimal n.pca for DAPC 

xval <- xvalDapc((X), grp$grp, n.pca.max = 90, training.set = 0.9, result = "groupMean", n.rep = 100)
#print(xval)

# Extract values from xval
n_pcs_tested <- as.numeric(names(xval$`Mean Successful Assignment by Number of PCs of PCA`))
mean_success  <- as.numeric(xval$`Mean Successful Assignment by Number of PCs of PCA`)
opt_pcs       <- as.numeric(xval$`Number of PCs Achieving Highest Mean Success`)

# Save plot of mean successful assignment vs number of PCs

plot(n_pcs_tested, mean_success,
     type = "b", pch = 19, col = "blue",
     xlab = "Number of PCs",
     ylab = "Mean Successful Assignment")
abline(v = opt_pcs, col = "red", lty = 2)
legend("bottomright",
       legend = paste("Optimal PCs =", opt_pcs),
       col = "red", lty = 2, bty = "n")



# Check the cluster assignments

opt_pcs <- as.numeric(xval$`Number of PCs Achieving Highest Mean Success`)
# opt_pcs = 10


# 6) DAPC using validated opt_pcs
# >>> on 2,122 individuals x 21 loci is one of the slowest lines in the file.
dapc1 <- dapc(data_GenInd, grp$grp, n.pca = opt_pcs, n.clust = 2, n.da = 1)

# >>> carries the same information as Fig. S3.
# save DAPC scatter plot

# Plot scatter
scatter(dapc1,
        posi.da = "bottomright",
        cex = 1.2,
        scree.da = FALSE,
        bg = "white")



# >>> listed in README.md. Membership probabilities are already shown in Fig. S3.

# Plot compoplot
compoplot(dapc1,
          txt.leg = paste("Cluster", 1:2),
          lab = "",
          ncol = 1,
          xlab = "Individuals",
          col = alpha(c("red","blue"), 0.8),
          border = NA)




# Compare assignments
table(grp$grp, dapc1$assign)


# Variance explained by retained PCs
pca_var <- dapc1$pca.eig # extract the PCA eigenvalues
sum(pca_var[1:10]) / sum(pca_var) * 100 # calculate the percentage of variance explained by the first 10 PCs

# EXPORT: the three values quoted in Methods 2.3 (10 PCs retained, 99% mean
# correct reassignment, 46% of the genetic variance) were printed only.
dapc_summary <- data.frame(
  optimal_n_pca           = opt_pcs,
  mean_correct_assignment = max(mean_success),
  variance_explained_pct  = sum(pca_var[1:opt_pcs]) / sum(pca_var) * 100,
  n_clusters              = 2
)
kable(dapc_summary, digits = 3)
write.csv(dapc_summary,
          file = "output/tables/Table_DAPC_crossvalidation_summary.csv",
          row.names = FALSE)

# Convert the posterior probabilities to a data frame
posterior_data <- as.data.frame(dapc1$posterior)

# Add individual id as a column in the data frame
posterior_data$Individual <- rownames(posterior_data)

# Melt the data frame for visualization
melted_data <- reshape2::melt(posterior_data, id.vars = "Individual", variable.name = "Cluster", value.name = "Probability")


# Scatter plot (show names of ind.)
cluster_assignments <- ggplot(melted_data, aes(x = Cluster, y = Probability, color = Cluster, label = Individual)) +
  geom_point() +
  xlab("Genotypic clusters") +
  ylab("Cluster probability") +
  labs(color = "Clusters") +
  theme_minimal(base_family = "serif", base_size = 12) +  
  theme(
    panel.background = element_rect(fill = "transparent", color = NA),  
    plot.background = element_rect(fill = "transparent", color = NA),   
    legend.background = element_rect(fill = "transparent"),             
    legend.key = element_rect(fill = "transparent")                     
  ) +
# The Fig. S3 caption says "with the names of individuals of lower probability",
# but this labelled all 2,122 individuals, which is the overlap the PCI data
# editor reported as unreadable. Only individuals that fall below the 0.8
# membership threshold in both clusters are labelled now.
geom_text_repel(
  data = subset(melted_data, Probability > 0.2 & Probability < 0.8),
  max.overlaps = Inf, size = 3, min.segment.length = 0,
  box.padding = 0.6, seed = 123) # to avoid overlapping labels
  

cluster_assignments

ggsave("output/figures/FigS3_cluster_assignments.png", cluster_assignments, width = 8, height = 6, dpi = 1200, bg = "white")


#reomve uncertin cluster
# Subset the main table to include only individuals with probability >= 0.8 in either cluster
certain_data <- genotype_mlp[posterior_data[, 1] >= 0.80 | posterior_data[, 2] >= 0.80 , ]

uncertain_data <- genotype_mlp[posterior_data[, 1] < 0.80 & posterior_data[, 2] < 0.80, ]

genotype_data$Cluster <- ifelse(posterior_data[, 1] >= 0.80, "Cluster 1", #80% 
                       ifelse(posterior_data[, 2] >= 0.80, "Cluster 2", 
                                            "None Determined"))


# Inspect the assignments
table(genotype_data$Cluster)

#Remove uncertain_data
genotype_data <- subset(genotype_data, Cluster %in% c("Cluster 1", "Cluster 2"))



```



### Population genetic indices per cluster - Table 1a


```{r}

# Creat it again because we removes uncertain data
# Select columns with "Mlp" in the name and the first column as isolate id
genotype_cols <- c(grep("Mlp", names(genotype_data), value = TRUE))
genotype_mlp <- genotype_data[, genotype_cols]

# Make isolate id as column names 
rownames(genotype_mlp) <- rownames(genotype_data)


# Convert to genind object
data_GenInd_Cluster <- df2genind(
  X = genotype_mlp,   
  sep = NULL,
  ncode = 3,           
  ind.names = rownames(genotype_mlp), 
  loc.names = colnames(genotype_mlp),  
  pop = genotype_data$Cluster,         
  NA.char = "999",                      
  ploidy = 2,
  type = "codom",                      
  strata = NULL,
  hierarchy = NULL,
  check.ploidy = getOption("adegenet.check.ploidy")
)

Nb_Pop = length(levels(data_GenInd_Cluster@pop))

# Assigning the vectors to be calculated
N <- vector(mode ="integer", length = Nb_Pop)
GsurN<- vector(mode ="numeric", length = Nb_Pop)
Ho <- vector(mode ="numeric", length = Nb_Pop)
Hs <- vector(mode ="numeric", length = Nb_Pop)
Fis <- vector(mode ="numeric", length = Nb_Pop)
rbarD<- vector(mode ="numeric", length = Nb_Pop)
Fis_sd <- vector(mode ="numeric", length = Nb_Pop)
Fis_var <- vector(mode ="numeric", length = Nb_Pop)


Pop <- levels(data_GenInd_Cluster$pop)

# Poppr to retrieve the first indices N, G/N and rbarD
Table_PPR <- poppr(data_GenInd_Cluster) 

rbarD <- Table_PPR$rbarD[1:Nb_Pop]
N <- Table_PPR$N[1:Nb_Pop]
GsurN <- (Table_PPR$MLG[1:Nb_Pop]-1)/(Table_PPR$N[1:Nb_Pop]-1)

# Calculation proba of LD 
ProbaLD <- vector(mode ="numeric", length = Nb_Pop)
for (i in 1:Nb_Pop) {
  Temp_Sample <- popsub(data_GenInd_Cluster, Table_PPR$Pop[i])
  ProbaLD[i] = ia(Temp_Sample, sample = 999, plot = F)[4]
}


# Use Fstat to calculate other Ho, He, Fis and Ar indices
data_Fstat <-genind2hierfstat(data_GenInd_Cluster)

# Loops to calculate the desired indices for each population
a =0
for (i in levels(data_Fstat$pop) ){
  Poptmp <- data_GenInd_Cluster[which(data_GenInd_Cluster$pop==i),]
  fstat_basic_Temporel <- basic.stats(Poptmp)
  a = a+1
  Ho[a]<-fstat_basic_Temporel$overall["Ho"]  
  Hs[a]<-fstat_basic_Temporel$overall["Hs"]  
  Fis[a]<-fstat_basic_Temporel$overall["Fis"] 
  Fis_sd[a]<- sd(fstat_basic_Temporel$perloc$Fis)
  Fis_var[a]<- var(fstat_basic_Temporel$perloc$Fis)

}

# Calculating allelic richness
Obj_Ar <- allelic.richness(data_Fstat[which(data_Fstat[,"pop"]!="NA"),])
Ar_per_loc <- Obj_Ar$Ar
Ar<- vector(mode ="numeric", length = Nb_Pop)
for (i in 1:Nb_Pop){ 
  Ar[i] = mean(Ar_per_loc[,i])
}


## Final table - genetic indices per cluster (manuscript Table 1)
MLG <- Table_PPR$MLG[1:Nb_Pop]
MLL <- sapply(Pop, function(p) {
  length(unique(genotype_data$organised_MLL[genotype_data$Cluster == p]))
})

# EXPORT: ProbaLD (computed above with ia(..., sample = 999)) is the permutation
# p-value of rd. It is the evidence behind "high linkage disequilibrium" in
# Results 3.2, it costs 999 permutations per cluster to compute, and it was
# being thrown away. It is added to Table 1 here.
Tab_Indices_per_pop <- rbind(N, MLG, MLL, GsurN, Ar, Ho, Hs, Fis, Fis_var, rbarD, ProbaLD)
colnames(Tab_Indices_per_pop) <- Pop
Tab_Indices_per_pop <- t(Tab_Indices_per_pop)

# Column names follow manuscript Table 1.
# Pareto beta is not computed by this script; add it from the manuscript
# methods if an exact match to the published Table 1 is required.
colnames(Tab_Indices_per_pop) <- c("N", "MLG", "MLL", "R", "Ar",
                                   "Ho", "He", "Fis", "Var (FIS)", "rd", "p (rd)")
kable(Tab_Indices_per_pop, digits = 3)

# Export Table 1 as CSV
write.csv(
  data.frame(Cluster = rownames(Tab_Indices_per_pop), Tab_Indices_per_pop,
             check.names = FALSE),
  file = "output/tables/Table1_a_genetic_indices_cluster_approach.csv",
  row.names = FALSE
)

```


### Assigning the reproductive mode from FIS


```{r}
# Tab_Indices_per_pop is the dataframe containing the results
# Loop through each row of the dataframe
# Define the  reproduction mode based on the Fis value (cluster approach)
for (i in 1:nrow(Tab_Indices_per_pop)) {
  cluster_name <- rownames(Tab_Indices_per_pop)[i]  # Get the cluster name
  
  # Check if the Fis value is negative
  if (Tab_Indices_per_pop[i, "Fis"] < 0) {
    # If Fis is negative, set to "Asex" 
    genotype_data$assignment[genotype_data$Cluster == cluster_name] <- "Asex"
  } else {
    # If Fis is non-negative, set the "Cluster" column to "Sex" for the corresponding cluster
    genotype_data$assignment[genotype_data$Cluster == cluster_name] <- "Sex"
  }
}

```

## Resampling approach

Depending on the MLL persistence, we can assign the reproduction mode to each lineage. If the same lineage exist at least two different years (to avoide clone mate) is asex, otherwise is sex.

### Multilocus lineages resampled across years
```{r}
# Add the Year column to the genotype_data table
genotype_data$Year <- general_data$Year[match(row.names(genotype_data), general_data$Isolate)]

# Calculate the frequency of each MLL-Year combination
repeated_MLL_table <- genotype_data [,c("organised_MLL","Year")]
rownames(repeated_MLL_table) <- NULL

# Rename the columns for clarity
colnames(repeated_MLL_table) <- c("organised_MLL",  "Year")



# Create an empty list to store the dictionary
mll_year_dict <- list()

# Loop through each row in the repeated_MLL_table
for (i in 1:nrow(repeated_MLL_table)) {
  mll <- repeated_MLL_table$organised_MLL[i]
  Year <- repeated_MLL_table$Year[i]
  
  # If the MLL is not yet in the dictionary, add it with the current year
  if (!(mll %in% names(mll_year_dict))) {
    mll_year_dict[[as.character(mll)]] <- list(Year)
  } else {
    # If the MLL is already in the dictionary, append the year to its list
    mll_year_dict[[as.character(mll)]] <- c(mll_year_dict[[as.character(mll)]], Year)
  }
}


# Convert the dictionary to a dataframe
mll_year_df <- data.frame(
  organised_MLL = as.integer(names(mll_year_dict)),
  Years = sapply(mll_year_dict, paste, collapse = ", ")
)


# Create a new column to store the counts
mll_year_df$YearCount <- NA

# Loop through each row in repeated_MLL_table
for (i in 1:nrow(mll_year_df)) {
  # Filter the mll_year_df for the current MLL
  filtered_df <- mll_year_df[mll_year_df$organised_MLL == mll_year_df$organised_MLL[i], ]
  
  # Count how many years are associated with the current MLL
  years_count <- length(unlist(strsplit(filtered_df$Years, ", ")))
  
  # Assign the count to the YearCount column
  mll_year_df$YearCount[i] <- years_count
}

# Loop through each row in repeated_MLL_table
for (i in 1:nrow(repeated_MLL_table)) {
  # Filter the mll_year_df for the current MLL
  filtered_df <- mll_year_df[mll_year_df$organised_MLL == repeated_MLL_table$organised_MLL[i], ]
  
  # Get unique years associated with the current MLL
  unique_years <- unique(unlist(strsplit(filtered_df$Years, ", ")))
  
  # Combine unique years into a single string
  years_string <- paste(unique_years, collapse = ", ")
  
  # Assign the combined years to the Years column in mll_year_df
  mll_year_df[mll_year_df$organised_MLL == repeated_MLL_table$organised_MLL[i], "Years"] <- years_string
}

#remove all non repeated mll  
mll_year_df <- mll_year_df[!(mll_year_df$YearCount == 1), ]

#Test
#mll_year_df [mll_year_df$organised_MLL == 1, ]

# Print the contingency table
# write.csv quotes the Years column (which contains commas), so the header
# and data rows always have a matching number of columns.
write.csv(mll_year_df, file = "output/tables/filtered_mll_years.csv", row.names = FALSE)
```


### Assigning the reproductive mode from lineage persistence

```{r}
for (i in 1:nrow(mll_year_df)) {
  if (mll_year_df$YearCount[i] >= 1) { 
    mll <- mll_year_df[mll_year_df$organised_MLL == mll_year_df$organised_MLL[i], ]
    years_string <- mll$Years
    years <- unlist(strsplit(years_string, ", "))
    num_years <- length(unique(years))  # Count unique years
    
    if (num_years >= 2) { # If the MLL exists in at least two years, assign "Asex"
      mll_year_df$Mll_persistence[i] <- "Asex"
    } else {
      mll_year_df$Mll_persistence[i] <- "Undetermined"
    }
  } else {
    mll_year_df$Mll_persistence[i] <- "Undetermined"
  }
}


# the following loop to assign the Mll_persistence to the genotype data table

for (i in 1:nrow(mll_year_df)) {
  mll_value <- mll_year_df$organised_MLL[i]
  Mll_persistence_value <- mll_year_df$Mll_persistence[i]
  
  genotype_data$Mll_persistence[genotype_data$organised_MLL == mll_value] <- Mll_persistence_value
}

genotype_data[is.na(genotype_data$Mll_persistence), "Mll_persistence"] <- "Undetermined"


```


### Population genetic indices per reproductive mode - Table 1b


```{r}
# Convert to genind object
final_GenInd <- df2genind(
  X = genotype_mlp,    
  sep = NULL,
  ncode = 3,           
  ind.names = rownames(genotype_mlp),  
  loc.names = colnames(genotype_mlp),  
  pop = genotype_data$Mll_persistence,                      
  NA.char = "999",                     
  ploidy = 2,
  type = "codom",                       
  strata = NULL,
  hierarchy = NULL,
  check.ploidy = getOption("adegenet.check.ploidy")
)


Nb_Pop = length(levels(final_GenInd@pop))

# Assigned the vectors to be
N <- vector(mode ="integer", length = Nb_Pop)
GsurN<- vector(mode ="numeric", length = Nb_Pop)
Ho <- vector(mode ="numeric", length = Nb_Pop)
Hs <- vector(mode ="numeric", length = Nb_Pop)
Fis <- vector(mode ="numeric", length = Nb_Pop)
rbarD<- vector(mode ="numeric", length = Nb_Pop)
Fis_sd <- vector(mode ="numeric", length = Nb_Pop)
Fis_var <- vector(mode ="numeric", length = Nb_Pop)

Pop <- levels(final_GenInd$pop)

# Poppr for N, G/N et rbarD
Table_PPR <- poppr(final_GenInd) # marche bien

rbarD <- Table_PPR$rbarD[1:Nb_Pop]
N <- Table_PPR$N[1:Nb_Pop]
GsurN <- (Table_PPR$MLG[1:Nb_Pop]-1)/(Table_PPR$N[1:Nb_Pop]-1)

# LD
ProbaLD <- vector(mode ="numeric", length = Nb_Pop)
for (i in 1:Nb_Pop) {
  Temp_Sample <- popsub(final_GenInd, Table_PPR$Pop[i])
  ProbaLD[i] = ia(Temp_Sample, sample = 999, plot = F)[4] # Attention c'est assez long à tourner 
}


# Ho, He, Fis et Ar
data_Fstat <-genind2hierfstat(final_GenInd)

# Per each population
a =0
for (i in levels(data_Fstat$pop) ){
  Poptmp <- final_GenInd[which(final_GenInd$pop==i),]
  fstat_basic_Temporel <- basic.stats(Poptmp)
  a = a+1
  Ho[a]<-fstat_basic_Temporel$overall["Ho"]  
  Hs[a]<-fstat_basic_Temporel$overall["Hs"]  
  Fis[a]<-fstat_basic_Temporel$overall["Fis"] 
  Fis_sd[a]<- sd(fstat_basic_Temporel$perloc$Fis)
  Fis_var[a]<- var(fstat_basic_Temporel$perloc$Fis)
}

# Ar
Obj_Ar <- allelic.richness(data_Fstat[which(data_Fstat[,"pop"]!="NA"),])
Ar_per_loc <- Obj_Ar$Ar
Ar<- vector(mode ="numeric", length = Nb_Pop)
for (i in 1:Nb_Pop){ 
  Ar[i] = mean(Ar_per_loc[,i])
}

## Final table - genetic indices per reproduction mode (resampling approach)
MLG <- Table_PPR$MLG[1:Nb_Pop]
MLL <- sapply(Pop, function(p) {
  length(unique(genotype_data$organised_MLL[genotype_data$Mll_persistence == p]))
})
# EXPORT: same as Table 1a - the LD permutation p-value (ProbaLD) is kept.
Tab_Indices_per_pop <- rbind(N, MLG, MLL, GsurN, Ar, Ho, Hs, Fis, Fis_var, rbarD, ProbaLD)
colnames(Tab_Indices_per_pop) <- Pop
Tab_Indices_per_pop <- t(Tab_Indices_per_pop)
colnames(Tab_Indices_per_pop) <- c("N", "MLG", "MLL", "R", "Ar",
                                   "Ho", "He", "Fis", "Var (FIS)", "rd", "p (rd)")
kable(Tab_Indices_per_pop, digits = 3)
write.csv(
  data.frame(Reproduction = rownames(Tab_Indices_per_pop), Tab_Indices_per_pop,
             check.names = FALSE),
  file = "output/tables/Table1_b_genetic_indices_resampling_approach.csv",
  row.names = FALSE
)

```




## Agreement between the two approaches - Table 2

```{r, echo=FALSE}
# Create the contingency table
contingency_table <- table(genotype_data$assignment, genotype_data$Mll_persistence)

# Print the contingency table
print(contingency_table)

# Rename the rows and columns of the contingency table.
# NOTE: these labels are positional. table() orders levels alphabetically, so
# this is correct only while the levels are ("Asex","Sex") and
# ("Asex","Undetermined"). If a third level ever appears the labels silently go
# wrong, so check print(contingency_table) above before trusting Table 2.
rownames(contingency_table) <- c("Asex_cluster", "Sex_cluster")
colnames(contingency_table) <- c("Asex_resampling", "Undetermined_resampling")

# Fisher's exact test
fisher_test <- fisher.test(contingency_table)
print(fisher_test)

# EXPORT: the p-value, odds ratio and confidence interval quoted in Results 3.2
# were printed to the console only.
fisher_clust_resamp <- data.frame(
  Test       = "Clustering vs resampling assignment",
  p_value    = fisher_test$p.value,
  odds_ratio = unname(fisher_test$estimate),
  CI_low     = fisher_test$conf.int[1],
  CI_high    = fisher_test$conf.int[2]
)
write.csv(fisher_clust_resamp,
          file = "output/tables/Table2_b_fisher_test_clustering_vs_resampling.csv",
          row.names = FALSE)

# Save Table 2: contingency table (clustering vs resampling approach)
write.csv(
  as.data.frame.matrix(contingency_table),
  file = "output/tables/Table2_contingency_clustering_vs_resampling.csv"
)

```


## Combined definition of the reproductive mode


```{r}

# If the individual is assigned to the Asex cluster or has Asex resampling, assign "Asex" to the Reproduction column. Otherwise, assign "sex".

genotype_data$Reproduction <- ifelse(genotype_data$assignment == "Asex" | genotype_data$Mll_persistence == "Asex", "Asex","Sex") 


```



### Population genetic indices per reproductive mode - Table 1c



```{r}
# Convert to genind object
final_GenInd <- df2genind(
  X = genotype_mlp,   
  sep = NULL,
  ncode = 3,          
  ind.names = rownames(genotype_mlp),  
  loc.names = colnames(genotype_mlp), 
  pop = genotype_data$Reproduction,                       
  NA.char = "999",                     
  ploidy = 2,
  type = "codom",                       
  strata = NULL,
  hierarchy = NULL,
  check.ploidy = getOption("adegenet.check.ploidy")
)


Nb_Pop = length(levels(final_GenInd@pop))


N <- vector(mode ="integer", length = Nb_Pop)
GsurN<- vector(mode ="numeric", length = Nb_Pop)
Ho <- vector(mode ="numeric", length = Nb_Pop)
Hs <- vector(mode ="numeric", length = Nb_Pop)
Fis <- vector(mode ="numeric", length = Nb_Pop)
rbarD<- vector(mode ="numeric", length = Nb_Pop)
Fis_sd <- vector(mode ="numeric", length = Nb_Pop)
Fis_var <- vector(mode ="numeric", length = Nb_Pop)

Pop <- levels(final_GenInd$pop)

# N, G/N et rbarD
Table_PPR <- poppr(final_GenInd) # marche bien

rbarD <- Table_PPR$rbarD[1:Nb_Pop]
N <- Table_PPR$N[1:Nb_Pop]
GsurN <- (Table_PPR$MLG[1:Nb_Pop]-1)/(Table_PPR$N[1:Nb_Pop]-1)

# LD
ProbaLD <- vector(mode ="numeric", length = Nb_Pop)
for (i in 1:Nb_Pop) {
  Temp_Sample <- popsub(final_GenInd, Table_PPR$Pop[i])
  ProbaLD[i] = ia(Temp_Sample, sample = 999, plot = F)[4] 
}


#Fstat 
data_Fstat <-genind2hierfstat(final_GenInd)

# Per each population
a =0
for (i in levels(data_Fstat$pop) ){
  Poptmp <- final_GenInd[which(final_GenInd$pop==i),]
  fstat_basic_Temporel <- basic.stats(Poptmp)
  a = a+1
  Ho[a]<-fstat_basic_Temporel$overall["Ho"]  
  Hs[a]<-fstat_basic_Temporel$overall["Hs"]  
  Fis[a]<-fstat_basic_Temporel$overall["Fis"] 
  Fis_sd[a]<- sd(fstat_basic_Temporel$perloc$Fis)
  Fis_var[a]<- var(fstat_basic_Temporel$perloc$Fis)
}

# Ar
Obj_Ar <- allelic.richness(data_Fstat[which(data_Fstat[,"pop"]!="NA"),])
Ar_per_loc <- Obj_Ar$Ar
Ar<- vector(mode ="numeric", length = Nb_Pop)
for (i in 1:Nb_Pop){ 
  Ar[i] = mean(Ar_per_loc[,i])
}

## Final table - genetic indices per reproduction mode (combined approach)
MLG <- Table_PPR$MLG[1:Nb_Pop]
MLL <- sapply(Pop, function(p) {
  length(unique(genotype_data$organised_MLL[genotype_data$Reproduction == p]))
})
# EXPORT: same as Table 1a - the LD permutation p-value (ProbaLD) is kept.
Tab_Indices_per_pop <- rbind(N, MLG, MLL, GsurN, Ar, Ho, Hs, Fis, Fis_var, rbarD, ProbaLD)
colnames(Tab_Indices_per_pop) <- Pop
Tab_Indices_per_pop <- t(Tab_Indices_per_pop)
colnames(Tab_Indices_per_pop) <- c("N", "MLG", "MLL", "R", "Ar",
                                   "Ho", "He", "Fis", "Var (FIS)", "rd", "p (rd)")
kable(Tab_Indices_per_pop, digits = 3)
write.csv(
  data.frame(Reproduction = rownames(Tab_Indices_per_pop), Tab_Indices_per_pop,
             check.names = FALSE),
  file = "output/tables/Table1_c_genetic_indices_combination_approaches.csv",
  row.names = FALSE
)


# Save the per-isolate result table
write.csv(genotype_data, file = "output/tables/new_genotype_data.csv")

#clean
#rm (Ar_per_loc, data_Fstat, fstat_basic_Temporel, Obj_Ar, Poptmp , Table_PPR, Temp_Sample)

```


# Genetic indices within reproductive groups

## Asexual group, lineage by lineage - Table 4

```{r, echo=FALSE}

# Select columns with "Asex"
Asex_data <- genotype_data[genotype_data$Reproduction == "Asex", ]

# Select columns with "Mlp" in the name and the first column as isolate id
genotype_cols_Asex <- c(grep("Mlp", names(Asex_data), value = TRUE))
genotype_Asex <- Asex_data[, genotype_cols_Asex]
rownames(genotype_Asex) <- rownames(Asex_data)



# Convert to genind object
data_GenInd_Asex_MLL <- df2genind(
  X = genotype_Asex,  
  sep = NULL,
  ncode = 3,           
  ind.names = rownames(genotype_Asex), 
  loc.names = colnames(genotype_Asex),  
  pop = Asex_data$organised_MLL,                      
  NA.char = "999",                      
  ploidy = 2,
  type = "codom",                      
  strata = NULL,
  hierarchy = NULL,
  check.ploidy = getOption("adegenet.check.ploidy")
)
data_Genclone_Asex <- as.genclone(data_GenInd_Asex_MLL)

# >>> the asexual set computes a full pairwise distance matrix and draws a plot
# >>> that appears in no figure. The MLL threshold was already fixed above.
#data_Asex_filtered <-filter_stats(data_Genclone_Asex, distance = diss.dist, plot = TRUE, missing = "asis")

Nb_Pop = length(levels(data_GenInd_Asex_MLL@pop))



N <- vector(mode ="integer", length = Nb_Pop)
GsurN<- vector(mode ="numeric", length = Nb_Pop)
Ho <- vector(mode ="numeric", length = Nb_Pop)
Hs <- vector(mode ="numeric", length = Nb_Pop)
Fis <- vector(mode ="numeric", length = Nb_Pop)
rbarD<- vector(mode ="numeric", length = Nb_Pop)
Fis_sd <- vector(mode ="numeric", length = Nb_Pop)
# Fis_var was assigned in the loop below without being declared here, so it kept
# the length of the previous chunk's vector. Declared now.
Fis_var <- vector(mode ="numeric", length = Nb_Pop)

Pop <- levels(data_GenInd_Asex_MLL$pop)

# N, G/N et rbarD
Table_PPR <- poppr(data_GenInd_Asex_MLL) 
rbarD <- Table_PPR$rbarD[1:Nb_Pop]
N <- Table_PPR$N[1:Nb_Pop]
GsurN <- (Table_PPR$MLG[1:Nb_Pop]-1)/(Table_PPR$N[1:Nb_Pop]-1)


# Ho, He, Fis et Ar

data_Fstat <-genind2hierfstat(data_GenInd_Asex_MLL)

# Per each population
a =0
for (i in levels(data_Fstat$pop) ){
  Poptmp <- data_GenInd_Asex_MLL[which(data_GenInd_Asex_MLL$pop==i),]
  fstat_basic_Temporel <- basic.stats(Poptmp)
  a = a+1
  Ho[a]<-fstat_basic_Temporel$overall["Ho"]  
  Hs[a]<-fstat_basic_Temporel$overall["Hs"]  
  Fis[a]<-fstat_basic_Temporel$overall["Fis"] 
  Fis_sd[a]<- sd(fstat_basic_Temporel$perloc$Fis)
  Fis_var[a]<- var(fstat_basic_Temporel$perloc$Fis)

}

# Ar
Obj_Ar <- allelic.richness(data_Fstat[which(data_Fstat[,"pop"]!="NA"),])
Ar_per_loc <- Obj_Ar$Ar
Ar<- vector(mode ="numeric", length = Nb_Pop)
for (i in 1:Nb_Pop){ 
  Ar[i] = mean(Ar_per_loc[,i])
}

# Final table - genetic indices per asexual MLL
Tab_Indices_per_pop <- rbind(N, GsurN, Ar, Ho, Hs, Fis, rbarD, Fis_sd, Fis_var)
colnames(Tab_Indices_per_pop) <- Pop
Tab_Indices_per_mll <- t(Tab_Indices_per_pop)
kable(head(Tab_Indices_per_mll, 10), digits = 3)

# Manuscript Table 4: characteristics of the seven most abundant asexual MLLs
mlg_per_mll <- aggregate(MLG ~ organised_MLL, data = Asex_data,
                         FUN = function(x) length(unique(x)))
year_range  <- aggregate(Year ~ organised_MLL, data = Asex_data,
                         FUN = function(y) paste(min(y), max(y), sep = ":"))

asex_mll_tab <- data.frame(
  MLL = as.integer(rownames(Tab_Indices_per_mll)),
  N   = Tab_Indices_per_mll[, "N"],
  R   = Tab_Indices_per_mll[, "GsurN"],
  AR  = Tab_Indices_per_mll[, "Ar"],
  HO  = Tab_Indices_per_mll[, "Ho"],
  HE  = Tab_Indices_per_mll[, "Hs"],
  check.names = FALSE
)
asex_mll_tab$MLG <- mlg_per_mll$MLG[match(asex_mll_tab$MLL, mlg_per_mll$organised_MLL)]
asex_mll_tab$`Year range` <- year_range$Year[match(asex_mll_tab$MLL,
                                                   year_range$organised_MLL)]

# Keep the seven most abundant lineages, columns ordered as in the manuscript
asex_mll_tab <- asex_mll_tab[order(asex_mll_tab$N, decreasing = TRUE), ]
asex_mll_tab <- head(asex_mll_tab[, c("MLL", "MLG", "N", "Year range",
                                      "R", "AR", "HO", "HE")], 7)

kable(asex_mll_tab, digits = 3, row.names = FALSE)
# Save full per-MLL table for all asexual lineages
write.csv(
  data.frame(MLL = rownames(Tab_Indices_per_mll), Tab_Indices_per_mll,
             check.names = FALSE),
  file = "output/tables/Table_asexual_MLL_genetic_indices.csv",
  row.names = FALSE
)
# Save Table 4: top 7 most abundant asexual MLLs
write.csv(asex_mll_tab,
          file = "output/tables/Table4_top7_asexual_MLLs.csv",
          row.names = FALSE)


```



## Sexual group, lineage by lineage

```{r, echo=FALSE}

# Select columns with "Sex" 
Sex_data <- genotype_data[genotype_data$Reproduction == "Sex", ]
genotype_cols_Sex <- c(grep("Mlp", names(Sex_data), value = TRUE))
genotype_Sex <- Sex_data[, genotype_cols_Sex]

rownames(genotype_Sex) <- rownames(Sex_data)



# Convert to genind object
data_GenInd_Sex_MLL <- df2genind(
  X = genotype_Sex,   
  sep = NULL,
  ncode = 3,           
  ind.names = rownames(genotype_Sex),  
  loc.names = colnames(genotype_Sex), 
  pop = Sex_data$organised_MLL,                       
  NA.char = "999",                      
  ploidy = 2,
  type = "codom",                      
  strata = NULL,
  hierarchy = NULL,
  check.ploidy = getOption("adegenet.check.ploidy")
)
data_Genclone_Sex<- as.genclone(data_GenInd_Sex_MLL)


#data_Sex_filtered <-filter_stats(data_Genclone_Sex, distance = diss.dist, plot = TRUE, missing = "asis")

Nb_Pop = length(levels(data_GenInd_Sex_MLL@pop))


N <- vector(mode ="integer", length = Nb_Pop)
GsurN<- vector(mode ="numeric", length = Nb_Pop)
Ho <- vector(mode ="numeric", length = Nb_Pop)
Hs <- vector(mode ="numeric", length = Nb_Pop)
Fis <- vector(mode ="numeric", length = Nb_Pop)
rbarD<- vector(mode ="numeric", length = Nb_Pop)
Fis_sd <- vector(mode ="numeric", length = Nb_Pop)
# Same as in the asexual block: Fis_var is used below but was never declared.
Fis_var <- vector(mode ="numeric", length = Nb_Pop)


Pop <- levels(data_GenInd_Sex_MLL$pop)

# N, G/N et rbarD
Table_PPR <- poppr(data_GenInd_Sex_MLL) 

rbarD <- Table_PPR$rbarD[1:Nb_Pop]
N <- Table_PPR$N[1:Nb_Pop]
GsurN <- (Table_PPR$MLG[1:Nb_Pop]-1)/(Table_PPR$N[1:Nb_Pop]-1)


# Ho, He, Fis et Ar

data_Fstat <-genind2hierfstat(data_GenInd_Sex_MLL)

# Per population
a =0
for (i in levels(data_Fstat$pop) ){
  Poptmp <- data_GenInd_Sex_MLL[which(data_GenInd_Sex_MLL$pop==i),]
  fstat_basic_Temporel <- basic.stats(Poptmp)
  a = a+1
  Ho[a]<-fstat_basic_Temporel$overall["Ho"]  
  Hs[a]<-fstat_basic_Temporel$overall["Hs"]  
  Fis[a]<-fstat_basic_Temporel$overall["Fis"] 
  Fis_sd[a]<- sd(fstat_basic_Temporel$perloc$Fis)
  Fis_var[a]<- var(fstat_basic_Temporel$perloc$Fis)

}

# Ar
Obj_Ar <- allelic.richness(data_Fstat[which(data_Fstat[,"pop"]!="NA"),])
Ar_per_loc <- Obj_Ar$Ar
Ar<- vector(mode ="numeric", length = Nb_Pop)
for (i in 1:Nb_Pop){ 
  Ar[i] = mean(Ar_per_loc[,i])
}

## Final table - genetic indices per sexual MLL
Tab_Indices_per_pop <- rbind(N, GsurN, Ar, Ho, Hs, Fis, rbarD, Fis_sd, Fis_var)
colnames(Tab_Indices_per_pop) <- Pop
Tab_Indices_per_mll <- t(Tab_Indices_per_pop)
kable(head(Tab_Indices_per_mll, 10), digits = 3)
write.csv(
  data.frame(MLL = rownames(Tab_Indices_per_mll), Tab_Indices_per_mll,
             check.names = FALSE),
  file = "output/tables/genetic_indices_per_sexual_MLL.csv",
  row.names = FALSE
)

# Assuming 'Sex_data' is your data frame containing MLG and MLL information
mlg_counts <- aggregate(MLG ~ organised_MLL, data = Sex_data, FUN = function(x) length(unique(x)))
rownames(mlg_counts) <- mlg_counts[,1]

#Sexual_lineages <- merge(mlg_counts,Tab_Indices_per_mll,by="row.names",all.x=TRUE)



rm (Ar_per_loc, data_Fstat, data_Genclone_Asex, data_Genclone_Sex,
    data_GenInd_Cluster, data_GenInd_Sex_MLL, fstat_basic_Temporel,mlg_counts, Obj_Ar, Poptmp , 
    Table_PPR, Temp_Sample)

```




# Genetic relationships among lineages (manuscript Results 3.4)

## Shared annotation table for Figures 3 and 4

Both figures describe the same individuals: Figure 3 uses them for the rings
around the tree, Figure 4 for the abundance bars. The table is therefore built
once here, before either figure, so the two chunks are independent of each
other.

```{r}
# Add the region to the genotype data
genotype_data$Region <- general_data$Region[match(row.names(genotype_data), general_data$Isolate)]

# One row per individual: lineage, reproductive mode, year and region.
dat1 <- data.frame(
  ID       = rownames(genotype_data),
  MLL      = genotype_data$organised_MLL,
  Group    = genotype_data$Reproduction,
  Year     = as.factor(genotype_data$Year),
  Location = genotype_data$Region
)
```


## Figure 3 - neighbour-joining tree

```{r}
##  Neighbour-joining tree with 1000 bootstrap replicates

# Convert to genind object
final_GenInd <- df2genind(
  X = genotype_mlp,  
  sep = NULL,
  ncode = 3,         
  ind.names = rownames(genotype_mlp),  
  loc.names = colnames(genotype_mlp),  
  pop = genotype_data$organised_MLL,                 
  NA.char = "999",                    
  ploidy = 2,
  type = "codom",                    
  strata = NULL,
  hierarchy = NULL,
  check.ploidy = getOption("adegenet.check.ploidy")
)

# Snapshot of the working objects, so the slow first half of the script does
# not have to be re-run when returning to the analysis: load("output/my_data.RData")
# It is written under output/ like every other product of the run, so it lands
# in the same place whether the script runs in RStudio, in a terminal, or in
# the Docker container (where only /project/output is mounted).
save(genotype_mlp, genotype_data, general_data, file = "output/my_data.RData")

#tree2 <- aboot(final_GenInd, distance = diss.dist, tree = "nj",
#           missing = "asis", sample = 1000, quiet = TRUE, threads = 20)

tree2 <- readRDS("tree_nj_boot1000.rds")

tree2$Nnode                                    # expect 2113
sort(unique(as.numeric(tree2$node.label)))[1:15]   # should NOT be multiples of 20
summary(as.numeric(tree2$node.label))



# Bootstrap support in the format ggtree needs
# Support in the format ggtree needs
boot_df <- data.frame(node = (Ntip(tree2) + 1):(Ntip(tree2) + tree2$Nnode),
                      bootstrap = as.numeric(tree2$node.label))
boot_df$bootstrap[is.na(boot_df$bootstrap)] <- 0

sum(boot_df$bootstrap >= 70)   # how many points will actually appear


# Create group information based on 'Reproduction' column, Sex and Asex
groupInfo <- split(rownames(genotype_data), genotype_data$Reproduction)

# Group the tree labels based on the group information
tree <- groupOTU(tree2, groupInfo)
# npg_colors is defined once in the package-loading chunk at the top.
reproduction_colors <- c("Asex" = npg_colors[8], "Sex" = npg_colors[2])

# Create circular plot
options(ignore.negative.edge=TRUE)
p <- ggtree(tree, aes(color = group), layout="circular") %<+% boot_df +
  scale_color_manual(name = "Reproductive mode", values = reproduction_colors) +
  ggnewscale::new_scale_color() +
  geom_nodepoint(aes(subset = bootstrap >= 70, color = "\u2265 70%"), size = 1.2) +
  scale_color_manual(name = "Bootstrap", values = c("\u2265 70%" = "black")) +
  theme(
    panel.background = element_rect(fill='transparent', color = NA),
    plot.background = element_rect(fill='transparent', color=NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.background = element_rect(fill='transparent', color = NA),
    legend.box.background = element_rect(fill='transparent', color = NA),
    legend.text = element_text(color = 'black', face = 'bold', size = 13),
    legend.title = element_text(color = 'black', face = 'bold', size = 15)
  )


#p 
# dat1 is built in the "Shared annotation table" chunk above.
# Filter dat1 to include MLLs
dat1_filtered <- dat1 %>% filter(MLL %in% c(1, 2, 3, 4, 5, 6,8,9))
dat1_filtered$MLL <- as.factor(dat1_filtered$MLL)
dat1_filtered$Year <- as.character(dat1_filtered$Year) 

# Manually define the color mapping
mll_color_mapping <- c(
  "1" = npg_colors[7],
  "2" = npg_colors[5],
  "3" = npg_colors[9],
  "4" = npg_colors[4],
  "6" = npg_colors[1],
  "5" = "grey" , #  it's sexual lineage
  "8" = npg_colors[3],  
  "9" = npg_colors[6]
)

p2 <- p +
  geom_fruit(
    data = dat1_filtered,
    geom = geom_col,
    mapping = aes(y = ID, x = Year, fill = MLL),  
    pwidth = 0.4, 
    offset = 0.01, 
    axis.params = list(
      axis = "x",  
      text.angle = -90,  
      text.size = 3.2,
      vjust = -0.2,
      line.alpha = 0,
    ),
    grid.params = list(color = "black", linetype = 5, size = 0.1, alpha = 0.7)  # Add grid lines
  ) +
  scale_fill_manual(values = mll_color_mapping, drop = FALSE) +
  theme(
    panel.background = element_rect(fill='transparent', color = NA),
    plot.background = element_rect(fill='transparent', color=NA), # transparent plot bg
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    legend.background = element_rect(fill='transparent', color = NA),
    legend.box.background = element_rect(fill='transparent', color = NA),
    legend.text = element_text(color = 'black', face = 'bold', size = 13), 
    legend.title = element_text(color = 'black', face = 'bold', size = 15),
    panel.border = element_blank()
  ) 

p2

# README.md lists Fig3a_tree_plot1.png but nothing ever wrote it - this is the
ggsave("output/figures/Fig3_tree_nj_plot.png",  p2, width = 12, height = 9,  dpi = 600, bg = "white")
```



# Spatial distribution of reproductive modes (manuscript Results 3.3)

## Figure 2 - geographical distribution of reproductive modes, 2009 and 2011


```{r , fig.width=10, fig.height=8}

# summarize data by population and location
genotype_data$Pop <- general_data$Pop[match(row.names(genotype_data), general_data$Isolate)]

genotype_data$Long <- general_data$Long[match(row.names(genotype_data), general_data$Isolate)]

genotype_data$Lat <- general_data$Lat[match(row.names(genotype_data), general_data$Isolate)]


# Filter the data to include only the years 2009 and 2011
pop_data <- genotype_data %>%
  filter(Year %in% c("2009","2011")) %>% 
  group_by(Pop, Long, Lat, Reproduction) %>%
  summarize(n = n(), .groups = "drop") %>%  
  group_by(Pop) %>%
  mutate(percent = n / sum(n)) %>%
  ungroup() %>%
  dplyr::select(Pop, Long, Lat, Reproduction, n, percent)

# Creates an xyz object for use with the function draw.pie
xyz <- make.xyz(pop_data$Long, pop_data$Lat, pop_data$percent, pop_data$Reproduction)

# Sample size per site, matched to the order that make.xyz produced
n_by_site <- pop_data %>% group_by(Long, Lat) %>% summarize(N = sum(n), .groups = "drop")
Nvec <- n_by_site$N[match(paste(round(xyz$x, 5), round(xyz$y, 5)),
                          paste(round(n_by_site$Long, 5), round(n_by_site$Lat, 5)))]
cat("sample size: min =", min(Nvec), " median =", median(Nvec), " max =", max(Nvec), "\n")

# Pie radius: area proportional to sample size, so radius follows sqrt(N)
R_max <- 0.45
rad <- R_max * sqrt(Nvec / max(Nvec))

# Create a named vector for profile colors
profile_color_mapping <- c("Asex" = npg_colors[8], "Sex" = npg_colors[2])
transparent_colors <- adjustcolor(profile_color_mapping, alpha.f = 0.7)

# The plot of the pie chart above the map
svglite::svglite("output/figures/Fig2_geographical_distribution_2009_2011.svg",
                 width = 18/2.54, height = 15/2.54, bg = "white", pointsize = 12)

par(mai = c(0, 0, 0, 0), bty = "n")
basemap(xlim = c(-4.3, 9.1), ylim = c(41, 51), bg = "white", axes = FALSE)
map("france", fill = FALSE, col = "black", add = TRUE)
final_map <- draw.pie(xyz$x, xyz$y, xyz$z, radius = rad, col = transparent_colors,
                      border = "grey30", lwd = 0.3)

# Reproductive mode legend, top left over the English Channel
legend(-6.1, 50.95, title = "Reproductive mode",
       legend = c("Asexual", "Sexual"), fill = transparent_colors,
       border = "grey30", bty = "n", cex = 1, ncol = 1,
       title.adj = 0, text.col = "black")

# Sample size legend, one block per circle so each one can be moved on its own
# Radii of the reference circles, same formula as the pies
r10 <- R_max * sqrt(10 / max(Nvec))
r30 <- R_max * sqrt(30 / max(Nvec))
r60 <- R_max * sqrt(60 / max(Nvec))

# Equal gap between the circle EDGES, not between their centres
gap <- 0.25          # gap size in degrees, change this to spread them out
y10 <- 46.00         # centre of the top circle, move the whole stack from here
y30 <- y10 - r10 - gap - r30
y60 <- y30 - r30 - gap - r60

# Sample size legend, one block per circle so each one can be moved on its own
text(-6.10, y10 + r10 + 0.55, "Sample size (n)", adj = c(0, 0.5), cex = 1, font = 1)

draw.pie(-5.60, y10, matrix(1), radius = r10,
         col = "grey88", border = "grey30", lwd = 0.3)
text(-4.95, y10, "10", adj = c(0, 0.5), cex = 1)

draw.pie(-5.60, y30, matrix(1), radius = r30,
         col = "grey88", border = "grey30", lwd = 0.3)
text(-4.95, y30, "30", adj = c(0, 0.5), cex = 1)

draw.pie(-5.60, y60, matrix(1), radius = r60,
         col = "grey88", border = "grey30", lwd = 0.3)
text(-4.95, y60, "60", adj = c(0, 0.5), cex = 1)


# Scale bar and north arrow, sizes are in inches so they never change with the extent
prettymapr::addscalebar(plotepsg = 4326, pos = "bottomright", style = "bar",
                        widthhint = 0.25, htin = 0.08, label.cex = 0.9,
                        padin = c(0.15, 0.15))

prettymapr::addnortharrow(pos = "topright", scale = 0.5, padin = c(0.20, 0.20))

# Europe inset with France in red, drawn last because par(fig) switches the coordinates
rect(-5.90, 41.05, -2.50, 43.30, col = "white", border = "black", lwd = 0.7)
fx <- grconvertX(c(-5.90, -2.50), from = "user", to = "ndc")
fy <- grconvertY(c(41.05, 43.30), from = "user", to = "ndc")
par(fig = c(fx, fy))
par(mai = c(0.02, 0.02, 0.02, 0.02))
par(new = TRUE)   # new = TRUE must be the last par call, mai and fig reset it and the page gets wiped
plot.new()
plot.window(xlim = c(-12, 32), ylim = c(34, 62), asp = 1/cos(48*pi/180))
map("world", fill = TRUE, col = "grey88", border = "grey55", lwd = 0.2, add = TRUE)
map("world", regions = "France", fill = TRUE, col = "red", border = "red", lwd = 0.3, add = TRUE)

dev.off()
```


## Binomial GLMM on latitude and longitude - Table 3

Reproductive mode as a function of latitude and longitude, fitted on the
2009 and 2011 countrywide surveys (Methods 2.3). Sampling year enters as a
random effect. This chunk writes manuscript Table 3 and Figure S4.

```{r, echo=FALSE}

#Add site information
genotype_data$Location <- general_data$Site[match(row.names(genotype_data), general_data$Isolate)]
genotype_data$Year <- general_data$Year[match(row.names(genotype_data), general_data$Isolate)]


# Assuming your data includes a 'year' column
# Extract relevant columns
genotype_data <- genotype_data %>%
  mutate(
    nbsex = ifelse(Reproduction == "Sex", 1, 0),
    nbAsex = ifelse(Reproduction == "Asex", 1, 0)
  )



# Group by Location, Long, Lat, and year
aggregated_data <- genotype_data %>%
  group_by(Location, Long, Lat, Year) %>%
  summarise(
    nbsex = sum(nbsex),
    nbAsex = sum(nbAsex)
  )

# Create the final table
final_table <- aggregated_data %>%
  select(Location, Long, Lat, Year, nbsex, nbAsex)
final_table$Long <- as.numeric(final_table$Long)
final_table$Lat <- as.numeric(final_table$Lat)
final_table$Year <- as.numeric(final_table$Year)




### Figure S4 - effect of latitude on the proportion of sexual reproduction


# Filter the data to include only the years 2009 and 2011
final_table_filtered <- final_table %>%
  filter(Year %in% c(2009, 2011))

# Convert Year to factor in the filtered data
final_table_filtered$Year <- as.factor(final_table_filtered$Year)

# Fixed-effect glm on the same 2009 + 2011 data, kept as a sanity check that
# the Lat and Long effects hold without the Year random effect. It is printed
# only - it produces no figure and no table. Manuscript Fig. S4 is the GLMM
# version built below.
# Fit the regression model with the filtered data
regression_filtered <- glm(formula = cbind(nbsex, nbAsex) ~ Lat + Long + Year,
                           data = final_table_filtered,
                           family = binomial)

# Display the summary of the regression model
summary(regression_filtered)

# The glm version of the Fig. S4 plot used to be built here and immediately
# overwritten by the GLMM version below (same object name, p_lat_filtered), and
# its ggsave was commented out, so it never produced a file. Removed.

#GLMM

# Fit the GLMM with Year as a random effect
glmm_model <- glmer(
  formula = cbind(nbsex, nbAsex) ~ Lat + Long + (1 | Year),
  data = final_table_filtered,
  family = binomial
)
#Summary of the GLMM
summary(glmm_model)

# Save Table 3: GLMM fixed-effect coefficients (Lat, Long → reproductive mode)
glmm_coef <- as.data.frame(summary(glmm_model)$coefficients)
glmm_coef <- cbind(Predictor = rownames(glmm_coef), glmm_coef)
rownames(glmm_coef) <- NULL
colnames(glmm_coef) <- c("Predictor", "Estimate", "Std.Error", "z.value", "p.value")
write.csv(glmm_coef,
          file = "output/tables/Table3_GLMM_binomial_regression.csv",
          row.names = FALSE)

# EXPORT: Results 3.3 quotes the Year random effect (variance = 13.44,
# sd = 3.666). It was only visible in summary(glmm_model) on screen.
glmm_ranef <- as.data.frame(lme4::VarCorr(glmm_model))
glmm_ranef <- data.frame(Group    = glmm_ranef$grp,
                         Variance = glmm_ranef$vcov,
                         SD       = glmm_ranef$sdcor)
kable(glmm_ranef, digits = 3)
write.csv(glmm_ranef,
          file = "output/tables/Table3_b_GLMM_random_effect_year.csv",
          row.names = FALSE)

# Create prediction data for Latitude, holding Long at median, with 2009 and 2011
prediction_data_filtered <- expand.grid(
  Lat = seq(min(final_table_filtered$Lat), max(final_table_filtered$Lat), length.out = 100),
  Long = median(final_table_filtered$Long),  # Fix Long at median
  Year = factor(c("2009", "2011"))  # Match factor levels from data
)

# Predict probabilities, including random effects
prediction_data_filtered$predicted_prob <- predict(
  glmm_model, 
  newdata = prediction_data_filtered, 
  type = "response",  # Probabilities on response scale
  re.form = NULL      # Include random effects of Year
)

# Plot the data points and GLMM prediction curve for Latitude
p_lat_filtered <- ggplot(final_table_filtered, aes(x = Lat, y = nbsex / (nbsex + nbAsex))) +
  # Data points (semi-transparent)
  geom_point(aes(size = nbsex + nbAsex), alpha = 0.5, shape = 16, color = "gray40") +
  # GLMM prediction curves (slightly more opaque for visibility)
  geom_line(data = prediction_data_filtered, 
            aes(x = Lat, y = predicted_prob, color = Year),
            linewidth = 1.2, alpha = 0.8) +
  # Labels and scales
  labs(
    x = "Latitude",
    y = "Proportion of sexual reproduction"
  ) +
  scale_color_manual(values = c("2009" = npg_colors[3], "2011" = npg_colors[5])) +
  scale_size_continuous(range = c(2, 8), name = "Sample Size") +
  
  # Theme settings for transparency and font
  theme_minimal(base_size = 12, base_family = "serif") +
  theme(
    text = element_text(family = "serif"),
    legend.position = "top",
    legend.title = element_blank(),
    plot.title = element_text(face = "bold", hjust = 0.5),
    panel.grid.minor = element_blank(),
    # Full transparency settings:
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent"),
    legend.key = element_rect(fill = "transparent")
  )



p_lat_filtered

# Save GLMM regression plot
ggsave("output/figures/FigS4_effect_of_latitude_glmm.png", plot = p_lat_filtered, width = 6, height = 4, dpi = 1200, bg = "white")

```



# Rate of clonality

## Input files for ClonEstiMate

```{r}
# Create table for minimum spanning tree (MST) software by selecting relevant columns
MST_data <- c("Reproduction", "Year", grep("Mlp", names(genotype_data), value = TRUE))
MST_data <- genotype_data[, MST_data]

# convert row name to col
MST_data <- MST_data %>%
  mutate(Isolate = rownames(MST_data)) %>%
  relocate(Isolate, .after = Year) %>%
  filter (Year %in% c(2009,2011)) # keep only 2009 and 2011 data



# remove rownames
rownames(MST_data) <- NULL

# Rename the second and third columns
colnames(MST_data)[2:3] <- c("gen", "tag")

# Function to split six-digit numbers into two columns
split_alleles <- function(x) {
  a <- as.numeric(substr(x, 1, 3))
  b <- as.numeric(substr(x, 4, 6))
  return(data.frame(a, b))
}

# Apply the split function to each Mlp_ column
split_columns <- lapply(MST_data[, -c(1:3)], split_alleles)

# Combine all split columns into a single data frame
split_data <- do.call(cbind, split_columns)

# Remove the column names of each second column in the split allele pairs
num_cols <- ncol(split_data)
for (i in seq(2, num_cols, by = 2)) {
  colnames(split_data)[i] <- ""
}


# Generate the 'ID' column as a unique identifier like 'C1', 'C2', 'C3', etc.
MST_data$ID <- paste0("C", seq_len(nrow(MST_data)))


# Create a new data frame with the desired format
MST_data <- data.frame(
  Pop = MST_data$Reproduction,
  gen = MST_data$gen,
  ID = MST_data$ID,
  tag = MST_data$tag,  # assuming tag is the same as ID
  color = "Red",  # assuming 'Red' is the color for all
  Ploidy = 2,  # assuming ploidy is 2
  split_data
)


# Delete second col name for each allele
colnames(MST_data)[grep("Var", colnames(MST_data))] <- ""



# Read the per-isolate results table
new_genotype_data <- read.csv("output/tables/new_genotype_data.csv", row.names = 1)


# Check the Reproduction status in new_genotype_data and update the color in MST_data
for (i in seq_len(nrow(MST_data))) {
  tag_name <- MST_data$tag[i]
  
  # Check if the tag_name exists in the row names of new_genotype_data
  if (tag_name %in% rownames(new_genotype_data)) {
    reproduction_status <- new_genotype_data[tag_name, "Reproduction"]
    
    # Update the color based on the Reproduction status
    if (reproduction_status == "Sex") {
      MST_data$color[i] <- "Blue"
    } else if (reproduction_status == "Asex") {
      MST_data$color[i] <- "Red"
    }
  }
}


# View the final MST data
head(MST_data)

# Remove any rows have missing data (999)


# Note: ClonEstiMate requires a specific column layout with paired unnamed allele
# columns. write.table keeps that structure because quote = FALSE and
# col.names = TRUE leave the empty second allele headers empty.
#
# The same table is written twice, on purpose:
#   .csv - archive copy. Standard, opens in Excel, and is the version listed in
#          README.md and deposited with the repository.
#   .txt - working copy, tab-delimited. This is the file to load into
#          ClonEstiMate, which reads plain text.
# Both hold exactly the same rows; only the separator and the extension differ.

clonestimate_base <- "output/tables/MST_data_mlp_pop_as_Reproduction_for_cloneEstimate"

# Archive copy (comma-separated)
write.table(MST_data,
            paste0(clonestimate_base, ".csv"),
            sep = ",", quote = FALSE, row.names = FALSE, col.names = TRUE)

# Working copy for the software (tab-separated)
write.table(MST_data,
            paste0(clonestimate_base, ".txt"),
            sep = "\t", quote = FALSE, row.names = FALSE, col.names = TRUE)
```



# Persistence and distribution of asexual lineages (manuscript Results 3.5)

## Figure 4 - abundance of asexual lineages, by region (A) and by year (B)

```{r}

# Manually define the color mapping
mll_color_mapping <- c(
  "1" = npg_colors[7], 
  "2" = npg_colors[5], 
  "3" = npg_colors[9], 
  "4" = npg_colors[4],
  "6" = npg_colors[1], 
  "Other MLLs " = "#848482",  
  "8" = npg_colors[3],  
  "9" = npg_colors[6] 
)




asexual_lineages_years <- dat1 %>%
  filter(Group == "Asex") %>% # Filter only Asex lineages
  mutate(MLL = ifelse(MLL %in% c(1,2,3,4,6,8,9), as.character(MLL), "Other MLLs ")) %>% 
  group_by(Year,MLL) %>% # Group by Location and MLL
  summarise(n = n(), .groups = "drop") # Summarize the counts


# Bar plot (Location and MLL)
mll_years <- ggplot(asexual_lineages_years, aes(x = Year, y = n, fill = as.factor(MLL))) +
  geom_bar(stat = "identity", position = "stack") +
  labs(
    x = "Year",
    y = "Abundance",
    fill = "MLL"
  ) +
  scale_fill_manual(values = mll_color_mapping) +
  theme(
    text = element_text(size = 12),
    axis.title = element_text(size = 15, face = "bold"), 
    axis.text = element_text(size = 14, color = "black"),  # Reduced from 15 for better balance
    axis.text.x = element_text(angle = 0, hjust = 1),  # Improved x-axis label readability
    panel.background = element_rect(fill = "transparent"),
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.grid = element_blank(),
    legend.background = element_rect(fill = "transparent")
  ) + guides(fill = "none")  # Quick legend removal

mll_years

ggsave("output/figures/Fig4B_asex_mll_Year.png", mll_years, width = 17, height = 8, dpi = 1200, bg = "white")
  



# Creat data for asexual lineages

asexual_lineages_locations <- dat1 %>%
  filter(Group == "Asex") %>% # Filter only Asex lineages
  mutate(MLL = ifelse(MLL %in% c(1,2,3,4,6,8,9), as.character(MLL), "Other MLLs ")) %>% 
  group_by(Location,MLL) %>% # Group by Location and MLL
  summarise(n = n(), .groups = "drop") # Summarize the counts



# Bar plot ( Location and MLL)
mll_locations <- ggplot(asexual_lineages_locations, aes(x = n, y = Location, fill = as.factor(MLL))) +
  geom_bar(stat = "identity", position = "stack") +
  labs(
    x = "Abundance",
    y = "Location",
    fill = "MLL"
  ) +
  scale_fill_manual(values = mll_color_mapping) +
  theme(
    text = element_text( size = 12),
    axis.title = element_text(size = 15, face = "bold"), 
    axis.text = element_text(size = 14, color = "black"),  # Slightly reduced from 15 for better fit
    legend.title = element_text(size = 14, face = "bold", margin = margin(r = 20)), 
    legend.text = element_text(size = 12),  # Removed bold from legend text for cleaner look
    legend.position = "top",
    legend.direction = "horizontal",
    legend.box = "horizontal",
    legend.spacing.x = unit(0.2, "cm"),  # Added spacing between legend items
    legend.key.size = unit(0.5, "cm"),  # Adjusted legend key size
    panel.background = element_rect(fill = "transparent"),
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.grid = element_blank(),
    legend.background = element_rect(fill = "transparent")
  ) +
  guides(fill = guide_legend(nrow = 1, override.aes = list(size = 4)))  # Forces single row

mll_locations

ggsave("output/figures/Fig4A_asex_mll_Locations.png", mll_locations, width = 17, height = 8, dpi = 1200, bg = "white")
  



# Add a title to the combined plot
combined_plot <- plot_grid(
  mll_locations, 
  mll_years,
  ncol = 1,
  labels = c("A", "B"),
  label_size = 15,
  align = "v"
) 

combined_plot

ggsave("output/figures/Fig4_asex_mll_Year_Locations.png", combined_plot, width = 10, height = 8, dpi = 1200, bg = "white")


```


## Fisher's exact tests - lineage by year and lineage by region

```{r}
asexual_lineages_years <- dat1 %>%
  filter(Group == "Asex") %>% # Filter only Asex lineages
  group_by(Year,MLL) %>% # Group by Location and MLL
  summarise(n = n(), .groups = "drop") # Summarize the counts



# Year and MLLs contingency table
contingency_table_year_mll <- asexual_lineages_years %>%
  pivot_wider(names_from = MLL, values_from = n, values_fill = list(n = 0)) # Reshape to wide format


print(contingency_table_year_mll)

# Fisher's exact test for each MLL
# Convert contingency table to matrix (excluding the Year column)
fisher_matrix_year_mll <- as.matrix(contingency_table_year_mll[, -1]) 

fisher_result_year_mll <- fisher.test(fisher_matrix_year_mll, simulate.p.value = TRUE, B = 1e5) # 100,000 simulations
print(fisher_result_year_mll)


asexual_lineages_locations <- dat1 %>%
  filter(Group == "Asex") %>% # Filter only Asex lineages
  group_by(Location,MLL) %>% # Group by Location and MLL
  summarise(n = n(), .groups = "drop") # Summarize the counts




# Locations and MLLs contingency table
contingency_table_location_mll <- asexual_lineages_locations %>%
  pivot_wider(names_from = MLL, values_from = n, values_fill = list(n = 0)) # Reshape to wide format

print(contingency_table_location_mll)

# Convert contingency table to matrix (excluding the Location column)
fisher_matrix_locations_mll <- as.matrix(contingency_table_location_mll[, -1])

fisher_result_locations_mll <- fisher.test(fisher_matrix_locations_mll, simulate.p.value = TRUE, B = 1e5) # 100,000 simulations
print(fisher_result_locations_mll)

# EXPORT: Results 3.5 quotes P = 1e-05 for both the temporal and the spatial
# test, but the spatial result was not even printed. Both are exported here,
# together with the two contingency tables they are computed from.
fisher_mll_tests <- data.frame(
  Test       = c("Asexual MLL x Year", "Asexual MLL x Region"),
  p_value    = c(fisher_result_year_mll$p.value,
                 fisher_result_locations_mll$p.value),
  replicates = 1e5
)
kable(fisher_mll_tests, digits = 6)
write.csv(fisher_mll_tests,
          file = "output/tables/Table_fisher_MLL_year_region.csv",
          row.names = FALSE)
write.csv(contingency_table_year_mll,
          file = "output/tables/Table_contingency_asexual_MLL_by_year.csv",
          row.names = FALSE)
write.csv(contingency_table_location_mll,
          file = "output/tables/Table_contingency_asexual_MLL_by_region.csv",
          row.names = FALSE)




```

# Supplementary analyses


## Table S2 - late-season samplings at Cavalaire-sur-Mer

```{r}
# EXPORT: Results 3.5 describes the February samplings at Cavalaire-sur-Mer
# (8 individuals in 2014, 40 in 2024), how many lineages they belong to, and
# whether each lineage is asexual or sexual. None of it was produced by the
# code, so none of those numbers could be reproduced from the repository.
# This chunk rebuilds them and writes them out.

cav_site <- general_data$Site[match(rownames(genotype_data), general_data$Isolate)]
cavalaire <- genotype_data[grepl("Cavalaire", cav_site, ignore.case = TRUE), ]

cavalaire_summary <- cavalaire %>%
  group_by(Year, organised_MLL, Reproduction) %>%
  summarise(N_individuals = n(), .groups = "drop") %>%
  arrange(Year, desc(N_individuals))

# Years in which each of these lineages was sampled anywhere in France, so the
# persistence statements in the text ("an older lineage sampled in 1993, 1996
# and 2004", "extensively resampled in 2020") can be checked directly.
cavalaire_summary$Years_sampled_overall <- sapply(
  cavalaire_summary$organised_MLL,
  function(m) paste(sort(unique(genotype_data$Year[genotype_data$organised_MLL == m])),
                    collapse = ", "))

# Number of distinct MLGs per lineage at this site
cavalaire_summary$N_MLG <- sapply(
  seq_len(nrow(cavalaire_summary)),
  function(i) length(unique(cavalaire$MLG[
    cavalaire$organised_MLL == cavalaire_summary$organised_MLL[i] &
    cavalaire$Year          == cavalaire_summary$Year[i]])))

cavalaire_summary <- cavalaire_summary[, c("Year", "organised_MLL", "Reproduction",
                                           "N_individuals", "N_MLG",
                                           "Years_sampled_overall")]
colnames(cavalaire_summary) <- c("Year", "MLL", "Reproduction", "N", "MLG",
                                 "Years sampled overall")

kable(cavalaire_summary)
write.csv(cavalaire_summary,
          file = "output/tables/Table_S2_cavalaire_february_samplings.csv",
          row.names = FALSE)

# One-line check against the text: total per year
print(table(cavalaire$Year, cavalaire$Reproduction))
```
